### 2022-11-18 materials-lisa/analysis/3_projection-across-operators/analysis.Rnw
try this model

%!TEX TS-program = xelatex
%!TEX encoding = UTF-8 Unicode

%% document settings
\documentclass[10pt]{article}
\usepackage{euler}
\usepackage{xunicode,xltxtra}
\usepackage{fontspec}

\usepackage[a4paper, margin = 1 in, left = .5 in, right = .5 in]{geometry}

\usepackage{setspace}
\usepackage[dvipsnames]{xcolor}

% fonts
	\defaultfontfeatures{Mapping=tex-text, Ligatures=TeX}
	\setromanfont[Mapping=tex-text, Numbers={Proportional}]{Linux Biolinum}
	\setsansfont[Scale=MatchLowercase,Mapping=tex-text]{Optima}
	\setmonofont[Scale=MatchLowercase]{Andale Mono}

\usepackage{amsmath}
\usepackage{amssymb}
% \usepackage{mathtools}
\usepackage{linguex}
\usepackage{booktabs}

\usepackage[colorlinks=true,citecolor=MidnightBlue,urlcolor=MidnightBlue,linkcolor=MidnightBlue]{hyperref}

\title{Effects on projectivity ratings by Embedding Operator and Trigger --- \newline Data Analysis}
\author{Lisa Hofmann}
\date{\today}

\begin{document}
\setkeys{Gin}{width=1.0\textwidth}
<<setup, include=FALSE, echo=FALSE>>=
library(knitr)
# library(highr)
hook_output = knit_hooks$get('output')
knit_hooks$set(output = function(x, options) {
  # this hook is used only when the linewidth option is not NULL
  if (!is.null(n <- options$linewidth)) {
    x = xfun::split_lines(x)
    # any lines wider than n should be wrapped
    if (any(nchar(x) > n)) x = strwrap(x, width = n)
    x = paste(x, collapse = '\n')
  }
  hook_output(x, options)
})

knitr::opts_chunk$set(echo=TRUE, prompt=TRUE, comment=NA, tidy.opts=list(width.cutoff=50), 
                      fig.path='figures/figures', fig.align = "center", out.width="\\linewidth")
options(tidy=TRUE, linewidth = 100)
@

\maketitle
\setcounter{tocdepth}{2}
\tableofcontents

\section{Introducing the dataset}
<<load-data, echo=FALSE, results='hide'>>=
# set working directory to directory of script
# this.dir <- dirname(rstudioapi::getSourceEditorContext()$path)
setwd("/Users/lisahofmann/Dropbox/Mein Mac (MBP-von-Lisa)/Documents/Academic/Research Projects/projective content/experiments-github/materials-lisa/analysis/3_projection-across-operators")

# load data from exps into list
df_list <- list(data_1q = read.csv("../../../results/main/1_projaiQ/data/data_preprocessed.csv", 
                    header=TRUE, sep=","),
                data_1n = read.csv("../../../results/main/2_projaiN/data/data_preprocessed.csv", 
                    header=TRUE, sep=","),
                data_1m = read.csv("../../../results/main/3_projaiM/data/data_preprocessed.csv", 
                    header=TRUE, sep=","),
                data_1c = read.csv("../../../results/main/4_projaiC/data/data_preprocessed.csv", 
                    header=TRUE, sep=","),
                data_2q = read.csv("../../../results/main/5_projaiQ/data/data_preprocessed.csv", 
                    header=TRUE, sep=","),
                data_2n = read.csv("../../../results/main/6_projaiN/data/data_preprocessed.csv", 
                    header=TRUE, sep=","),
                data_2m = read.csv("../../../results/main/7_projaiM/data/data_preprocessed.csv", 
                    header=TRUE, sep=","),
                data_2c = read.csv("../../../results/main/8_projaiC/data/data_preprocessed.csv", 
                    header=TRUE, sep=","),
                data_3q = read.csv("../../../results/main/9_projaiQ/data/data_preprocessed.csv", 
                    header=TRUE, sep=","),
                data_3n = read.csv("../../../results/main/10_projaiN/data/data_preprocessed.csv", 
                    header=TRUE, sep=","),
                data_3m = read.csv("../../../results/main/11_projaiM/data/data_preprocessed.csv", 
                    header=TRUE, sep=","),
                data_3c = read.csv("../../../results/main/12_projaiC/data/data_preprocessed.csv", 
                    header=TRUE, sep=",")
                )

@

<<clean-data, echo=FALSE, results='hide', warning=FALSE, message=FALSE>>=
require(tidyverse)
library(stringr)
library(rlist)

# spread responses over separate columns for projectivity and at-issueness
# add label for which position at block appeared in
df_list <- lapply(df_list, function(df) {
  df = df %>%
    mutate(ai_block = ifelse(question_type == "ai", 
                             ifelse(block == "block1", "block1", "block2"), 
                             ifelse(block == "block1", "block2", "block1"))) %>%
    select(workerid,content,short_trigger,question_type,response,ai_block) %>%
    spread(question_type,response)
  df
})

# change cd verb names to match veridicality names
df_list <- lapply(df_list, function(df) {
  df = df %>%
    mutate(verb=recode(short_trigger, control = "MC", annoyed = "be_annoyed", 
                       be_right_that = "be_right", inform_Sam = "inform"))
  df
})

# add labels for operator, and experiment block
df_list = map2(df_list, names(df_list), function(df, name) {
  operator <- str_sub(name, -1, -1)
  exp_block <- str_sub(name, -2, -2)
  df$op <- rep(operator, length(df[,1]))
  df$exp_block <- rep(exp_block, length(df[,1]))
  return(df)
})

# recode relevant vectors as factors
df_list <- lapply(df_list, function(df) {
  df$content <- as.factor(df$content)
  df$verb <- as.factor(df$verb)
  df$workerid <- as.factor(df$workerid)
  df$op <- as.factor(df$op)
  df$exp_block <- as.factor(df$exp_block)
  df
})

# remove data from MC controls
df_list <- lapply(df_list, function(df) {
  df = filter(df, verb != "MC")
  df$verb = droplevels(df$verb)
  df$content = droplevels(df$content)
  df
})

# combine data from all experiments
data_all <- list.rbind(df_list)
@

<<data-overview>>=
str(data_all)
length(levels(data_all$workerid))
@

\noindent The dataset consists of $57160$ observations from $2682$ participants (recruited on the online platforms Prolific and Amazon Mechanical Turk), across 12 experiments.

We are interested in how highly participants rate speaker commitment to the truth of an embedded complement clause, coded as \texttt{projective} on a real-numbered sliding scale between $0-1$.\\

\noindent The complement clause was embedded under an attitude verb, which in turn was embedded under an entailment-cancelling operator. Our fixed effects factors manipulate the following:

\begin{enumerate}
  \item The choice of attitude verb (coded as \texttt{verb})
  \item The entailment-cancelling operator (coded as \texttt{op})
\end{enumerate}

\noindent The levels for our fixed effects factors are the following:
<<fixed-efcts-lvls>>=
levels(data_all$verb)
length(levels(data_all$verb))
levels(data_all$op)
length(levels(data_all$op))
@

\noindent We are interested in the effect on \texttt{projective} of \texttt{verb} and \texttt{op}, as well as their interaction, corresponding to a $20 \times 4$ factorial design, yielding
<<ncond>>=
length(levels(data_all$verb))*length(levels(data_all$op))
@
conditions.\\

\noindent We have $20$ items, corresponding to the content of the complement clause.
<<nitems>>=
levels(data_all$content)
length(levels(data_all$content))
@

\noindent We have roughly $36$ observations by item and condition. This is an approximate number, because the \texttt{op} manipulation is a between-studies manipulation, and the number of participants differs by experiment:
<<nobs>>=
# n observations
length(data_all[,1])

# observations by item
length(data_all[,1])/length(levels(data_all$content))
table(data_all$content)

# observations by verb
length(data_all[,1])/length(levels(data_all$verb))
table(data_all$verb)

# observations by operator
length(data_all[,1])/length(levels(data_all$op))
table(data_all$op)

# observations by item and condition
length(data_all[,1])/length(levels(data_all$content))/
  (length(levels(data_all$verb))*length(levels(data_all$op)))

@

\newpage
\section{Data Overview and Statistical Summaries}
\subsection{Distribution of projectivity ratings by operator:}
<<project-by-op-distr, echo = FALSE>>=
library(ggplot2)
data_all %>% 
  # mutate(verb = fct_reorder(cond, as.numeric(dataf$response), .fun = 'mean')) %>%
  ggplot(aes(x = projective, group = op)) +
  facet_grid(rows = vars(op)) +
  geom_histogram(binwidth = .01, fill = "lightblue") +
  # coord_flip() +
  xlab("") +
  ylab("") +
  labs(title = "")+
  theme_bw()
  # scale_fill_brewer(palette = "Blues")
@

\begin{itemize}
  \item These definitely do not look normal
  \item Maybe a beta-regression would be useful?
  \item But even that would be relying on some simplifying assumptions, since we might be ignoring the little bump in the middle 
\end{itemize}

\newpage
\subsection{Distribution of projectivity ratings by verb:}
<<project-by-v-distr, echo = FALSE>>=
data_all %>% 
  # mutate(verb = fct_reorder(cond, as.numeric(dataf$response), .fun = 'mean')) %>%
  ggplot(aes(x = projective, group = verb)) +
  facet_wrap(vars(verb)) +
  geom_histogram(binwidth = .01, fill = "lightblue") +
  # coord_flip() +
  xlab("") +
  ylab("") +
  labs(title = "")+
  theme_bw()
  # scale_fill_brewer(palette = "Blues")
@

\begin{itemize}
  \item Some of these also show a higher mass around the middle of the scale
  \item but it looks the beta-distribution could be a useful approximation
\end{itemize}

\newpage
\subsection{Means and confidence intervals for projectivity rating by operator}
<<summary-op, echo = FALSE, message=FALSE, cache=TRUE, out.width=".6\\textwidth">>=
# load helper functions
source('../../../results/main/helpers.R')

proj_means_op = data_all %>% group_by(op) %>%
  summarize(Mean = mean(projective), CILow = ci.low(projective), 
            CIHigh = ci.high(projective)) %>%
  mutate(YMin = Mean - CILow, YMax = Mean + CIHigh, 
         op = fct_reorder(as.factor(op),Mean)) %>% ungroup()

proj_means_op %>% mutate(op = fct_reorder(op, Mean, 
                                         .fun = mean)) %>% 
  ggplot(aes(x = op, y=Mean, color = op)) +
  # coord_cartesian(ylim = c(0,1)) +
  geom_point(aes(shape = op), size = 4) + 
  scale_shape_manual(values = c("M", "N", "Q", "C")) +
  geom_errorbar(aes(ymin=YMin,ymax=YMax), width=0.1) +
  geom_line() + 
  labs(title = "Mean projectivity by operator")+
  theme_bw() +
  scale_color_brewer(palette = "PRGn")

proj_means_op %>% mutate(op = fct_reorder(op, Mean, 
                                         .fun = mean)) %>% 
  ggplot(aes(x = op, y=Mean, color = op, group = verb)) +
  # coord_cartesian(ylim = c(0,1)) +
  geom_point(aes(shape = op), size = 4) + 
  scale_shape_manual(values = c("M", "N", "Q", "C")) +
  geom_errorbar(aes(ymin=YMin,ymax=YMax), width=0.1) +
  facet_wrap(vars(verb)) +
  geom_line() + 
  labs(title = "Mean projectivity by operator")+
  theme_bw() +
  scale_color_brewer(palette = "PRGn")

@
\noindent The following generalizations emerge:
\begin{itemize}
  \item Conditionals have the highest projectivity ratings
  \item Projectivity ratings for questions are higher than those for modals and negation, but lower than those for conditionals
  \item Modals and negation have the lowest projectivity ratings
  \item The ratings for negation look a little higher than for modals, but error bars overlap
\end{itemize}

\noindent Although these differences appear to be significant, they are quite small.\\

\newpage
\subsection{Means and confidence intervals for projectivity rating by verb:}
<<summary-v, echo = FALSE, message=FALSE, cache=TRUE, fig.width=14, fig.height=7>>=

proj_means_v = data_all %>% group_by(verb) %>%
  summarize(Mean = mean(projective), CILow = ci.low(projective), 
            CIHigh = ci.high(projective)) %>%
  mutate(YMin = Mean - CILow, YMax = Mean + CIHigh, 
         verb = fct_reorder(as.factor(verb),Mean)) %>% ungroup()

proj_means_v %>% mutate(verb = fct_reorder(verb, Mean, 
                                         .fun = mean)) %>% 
  ggplot(aes(x=verb, y=Mean)) +
  coord_cartesian(ylim=c(0,1)) +
  geom_point(color = "blue") +
  geom_errorbar(aes(ymin=YMin,ymax=YMax), width=0.1, color = "blue") +
  geom_line() + 
  labs(title = "Mean projectivity by predicate")+
  theme_bw()
@
\begin{itemize}
  \item We see gradual differences in projectivity between verbs
\end{itemize}

\subsection{Means and confidence intervals for projectivity rating by verb and operator:}
<<summary-combined, echo = FALSE, message=FALSE, cache=TRUE, fig.width=14, fig.height=7>>=
proj_means = data_all %>% group_by(verb, op) %>%
  summarize(Mean = mean(projective), CILow = ci.low(projective), 
            CIHigh = ci.high(projective)) %>%
  mutate(YMin = Mean - CILow, YMax = Mean + CIHigh, 
         verb = fct_reorder(as.factor(verb),Mean)) %>% ungroup()

# kable(proj_means, "latex", longtable = TRUE)

proj_means %>% mutate(verb = fct_reorder(verb, Mean, 
                                         .fun = mean)) %>% 
  mutate(op = fct_reorder(op, Mean, .fun = mean)) %>% 
  ggplot(aes(x=verb, y=Mean, group = op, color = op)) +
  coord_cartesian(ylim=c(0,1)) +
  geom_point(aes(shape = op), size = 4) + 
  scale_shape_manual(values = c("M", "N", "Q", "C")) +
  geom_errorbar(aes(ymin=YMin,ymax=YMax),width=0.1) +
  geom_line() + 
  labs(title = "Mean projectivity for predicate by operator")+
  theme_bw() +
  scale_color_brewer(palette = "PRGn")

@

\begin{itemize}
  \item We see interactions between verb and operator
  \item However, we do not see any group of verbs that could be characterized as \lq semi-factive\rq\ in the sense of Karttunen. Specifically, \emph{discover} does not follow the predicted pattern: It is not more projective under negation, but most projective in conditionals and questions.

  \item Two verbs show highest projectivity under negation: the anti-veridical \emph{pretend}, and the non-veridical \emph{think}. These are verb with relatively low overall projectivity.

  \item What else\dots?
\end{itemize}

\section{Analysis}

We are interested in how the response (projectivity ratings) depends on the verb and embedding operator.

<<>>=
# main analysis 4: predict projection from not-at-issueness, predicate, and their interaction ----

t = data_all %>% 
  mutate(verb = fct_relevel(verb, "suggest"))

m.dummy = lm(projective ~ ai*short_trigger, data=t)
summary(m.dummy)

# reference level set to "prove"

# model with short_trigger slope did not converge
m.4 = lmer(projective ~ cai*short_trigger + (1+cai|content) + (1+cai|workerid), 
           data = t, REML=F,control = lmerControl(optimizer ='optimx', optCtrl=list(method='L-BFGS-B')))
saveRDS(m.4, "../models/m.4.rds")
m.4 <- readRDS("../models/m.4.rds")
summary(m.4)
texreg(m.4,single.row = TRUE, leading.zero = FALSE, booktabs = TRUE)

@

<<>>=
formula <- response ~ op * verb
# formula <- response ~ op * verb + (1 | workerid) + (1 | item)
# formula <- response ~ op * verb + (1 + op + verb | workerid) + (1 + op + verb | item)
@

Trying a beta-regression. Our response variable has observations from a sliding scale between $(0,1)$. For beta regression, we rescale it to $[0,1]$, using method used in Degen \& Tonhauser (2022), from Smithson \& Verkuilen (2006).

<<>>=
data_all$response = (data_all$projective*(nrow(data_all)-1) + .5)/nrow(data_all)
@

<<>>=
# library(rstan)
# options(mc.cores=parallel::detectCores())
# rstan_options(auto_write=TRUE)
# library(brms)
# 
# model1 <- brm(formula = bf(response ~ op * verb ,
#                            phi ~ op * verb,
#                            family = Beta()),
#               family = Beta(),
#               data = data_all,
#               iter=10000, warmup=2000,chains=6)

@

\end{document}