### 2022-12-05 materials-lisa/analysis/3_projection-across-operators/data-analysis.Rnw
removed tex gyre termes font

%!TEX TS-program = xelatex
%!TEX encoding = UTF-8 Unicode

%% document settings
\documentclass[10pt]{article}
\usepackage{euler}
\usepackage{xunicode,xltxtra}
\usepackage{fontspec}

% fonts
	\defaultfontfeatures{Mapping=tex-text, Ligatures=TeX}
	\setromanfont[Mapping=tex-text, Numbers={Proportional}]{Linux Biolinum}
	\setsansfont[Scale=MatchLowercase,Mapping=tex-text]{Optima}
	\setmonofont[Scale=MatchLowercase]{Andale Mono}

\usepackage[a4paper, margin = 1 in, left = .5 in, right = .5 in]{geometry}

\usepackage{setspace}
\usepackage[dvipsnames]{xcolor}

\usepackage{amsmath}
\usepackage{amssymb}
% \usepackage{mathtools}
\usepackage{linguex}
\usepackage{booktabs}

\usepackage[colorlinks=true,citecolor=MidnightBlue,urlcolor=MidnightBlue,linkcolor=MidnightBlue]{hyperref}

\title{Effects on projectivity ratings by Embedding Operator and Trigger --- \newline Data Analysis}
\author{Hofmann, deMarneffe, Tonhauser}
\date{\today}

\begin{document}
\setkeys{Gin}{width=1.0\textwidth}

<<setup, include=FALSE, echo=FALSE>>=
library(knitr)
hook_output = knit_hooks$get('output')
knit_hooks$set(output = function(x, options) {
  # this hook is used only when the linewidth option is not NULL
  if (!is.null(n <- options$linewidth)) {
    x = xfun::split_lines(x)
    # any lines wider than n should be wrapped
    if (any(nchar(x) > n)) x = strwrap(x, width = n)
    x = paste(x, collapse = '\n')
  }
  hook_output(x, options)
})

knitr::opts_chunk$set(echo=TRUE, prompt=TRUE, comment=NA, tidy.opts=list(width.cutoff=50), 
                      fig.path='figures/', fig.align = "center", out.width="\\linewidth")
options(tidy=TRUE, linewidth = 100)
@

\maketitle
\setcounter{tocdepth}{2}
\tableofcontents

\section{Introducing the dataset}
<<load-data, echo=FALSE, results='hide'>>=
# set working directory to directory of script
# this.dir <- dirname(rstudioapi::getSourceEditorContext()$path)
# setwd(this.dir)
setwd("/Users/lisa/Documents/research/GPP-experiments/materials-lisa/analysis/3_projection-across-operators")

# load data
df <- read.csv("../../../results/main/meta-analyses/projectivity/data/data_combined.csv", 
                    header=TRUE, sep=",")

@

<<data-overview>>=
str(df)
df$workerid <- as.factor(df$workerid)
length(levels(df$workerid))
@

\noindent The dataset consists of $57160$ observations from $2682$ participants (recruited on the online platforms Prolific and Amazon Mechanical Turk), across 12 experiments.

We are interested in how highly participants rate speaker commitment to the truth of an embedded complement clause, coded as \texttt{projective} on a real-numbered sliding scale between $0-1$.\\

\noindent The complement clause was embedded under an attitude verb, which in turn was embedded under an entailment-cancelling operator. Our fixed effects factors manipulate the following:

\begin{enumerate}
  \item The choice of attitude verb (coded as \texttt{verb})
  \item The entailment-cancelling operator (coded as \texttt{op})
\end{enumerate}

\noindent The levels for our fixed effects factors are the following:
<<fixed-efcts-lvls>>=
df$verb <- as.factor(df$verb)
levels(df$verb)
length(levels(df$verb))
df$op <- as.factor(df$op)
levels(df$op)
length(levels(df$op))
@

\noindent We are interested in the effect on \texttt{projective} of \texttt{verb} and \texttt{op}, as well as their interaction, corresponding to a $20 \times 4$ factorial design, yielding
<<ncond>>=
length(levels(df$verb))*length(levels(df$op))
@
conditions.\\

\noindent We have $20$ items, corresponding to the content of the complement clause.
<<nitems>>=
df$content <- as.factor(df$content)
levels(df$content)
length(levels(df$content))
@

\noindent We have roughly $36$ observations by item and condition. This is an approximate number, because the \texttt{op} manipulation is a between-studies manipulation, and the number of participants differs by experiment:
<<nobs>>=
# n observations
length(df[,1])

# observations by item
length(df[,1])/length(levels(df$content))
table(df$content)

# observations by verb
length(df[,1])/length(levels(df$verb))
table(df$verb)

# observations by operator
length(df[,1])/length(levels(df$op))
table(df$op)

# observations by item and condition
length(df[,1])/length(levels(df$content))/
  (length(levels(df$verb))*length(levels(df$op)))

@

\newpage
\section{Statistical Summaries and Graphs}
\subsection{Projectivity rating by operator}

Distribution of projectivity ratings by operator with means and $95\%$ bootstrapped confidence intervals.

<<summary-op, echo = FALSE, message=FALSE, cache=TRUE, out.width=".6\\textwidth">>=
# packages for data manipulation and plots
require(ggplot2)
require(dplyr)
require(forcats)

# load helper functions
source('../../../results/main/helpers.R')

pmeans_op = df %>% group_by(op) %>%
  summarize(Mean = mean(projective), CILow = ci.low(projective),
            CIHigh = ci.high(projective)) %>%
  mutate(YMin = Mean - CILow, YMax = Mean + CIHigh,
         op = fct_reorder(as.factor(op),Mean)) %>% ungroup()

pmeans_op %>% mutate(op = fct_reorder(op, Mean,
                                          .fun = mean)) %>%
  ggplot(aes(x = op, y=Mean)) +
  geom_violin(data=data.frame(op=df$op, Mean=df$projective),scale="width",color="gray80") +
  geom_point(size = 1, color = "lightblue") +
  geom_errorbar(aes(ymin=YMin,ymax=YMax), width=0.1, color = "lightblue") +
  geom_line(color = "lightblue") +
  # labs(title = "Projectivity by operator")+
  theme_bw()

@

\noindent The following generalizations emerge:
\begin{itemize}
  \item Conditionals have the highest projectivity ratings
  \item Projectivity ratings for questions are higher than those for modals and negation, but lower than those for conditionals
  \item Modals and negation have the lowest projectivity ratings
  \item The ratings for negation look a little higher than for modals, but error bars overlap
\end{itemize}

\noindent Although these differences appear to be significant, they are quite small.\\

\newpage
\subsection{Distributions of projectivity rating by verb with means and 95\% bootstrapped confidence interval:}

<<summary-verb, echo = FALSE, message=FALSE, cache=TRUE, fig.width=8, fig.height=6>>=

pmeans_verb = df %>% group_by(verb) %>%
  summarize(Mean = mean(projective), CILow = ci.low(projective), 
            CIHigh = ci.high(projective)) %>%
  mutate(YMin = Mean - CILow, YMax = Mean + CIHigh, 
         verb = fct_reorder(as.factor(verb),Mean))

# plot of projectivity means, 95% CIs and participants' ratings 
pmeans_verb %>% 
  ggplot(aes(x=fct_reorder(verb, Mean), y=Mean)) +
  geom_violin(data=data.frame(verb=df$verb, Mean=df$projective),scale="width",color="gray80") +
  geom_errorbar(aes(ymin=YMin,ymax=YMax),width=0.3,color="black") +
  geom_point(size=0.5,color="black") +
  scale_y_continuous(limits = c(0,1),breaks = c(0,0.2,0.4,0.6,0.8,1.0)) +
  theme(axis.text.x = element_text(size = 10, angle = 45, hjust = 1)) +
  xlab("") + ylab("")

@

\begin{itemize}
  \item We see gradual differences in projectivity between verbs
\end{itemize}

\subsection{Distributions, means, and CIs for projectivity ratings by verb and operator:}
<<summary-combined-1>>=
pmeans = df %>% group_by(verb, op) %>%
  summarize(Mean = mean(projective), CILow = ci.low(projective),
            CIHigh = ci.high(projective)) %>%
  mutate(YMin = Mean - CILow, YMax = Mean + CIHigh,
         verb = fct_reorder(as.factor(verb),Mean)) %>% ungroup()

ggplot(pmeans, aes(x=verb, y=Mean)) +
  geom_violin(data=data.frame(verb=df$verb, Mean=df$projective),scale="width",color="gray80") +
  geom_errorbar(aes(ymin=YMin,ymax=YMax),width = 0.5, color="black") +
  geom_point(size=0.5,color="black") +
  facet_wrap(~op, labeller = labeller(op = c("m" = "Modal", "n" = "Negation", "q" = "Question", "c" = "Conditional"))) +
  scale_y_continuous(limits = c(0,1),breaks = c(0,0.2,0.4,0.6,0.8,1.0)) +
  theme(axis.text.x = element_text(size = 8, angle = 45, hjust = 1)) +
  ylab("Mean certainty rating") +
  xlab("Predicate")
@

\subsection{Means and confidence intervals for projectivity rating by verb and operator:}
<<summary-combined-2, echo = FALSE, message=FALSE, cache=TRUE, fig.width=14, fig.height=7>>=

pmeans %>% mutate(verb = fct_reorder(verb, Mean,
                                         .fun = mean)) %>%
  mutate(op = fct_reorder(op, Mean, .fun = mean)) %>%
  ggplot(aes(x=verb, y=Mean, group = op, color = op)) +
  coord_cartesian(ylim=c(0,1)) +
  geom_point(aes(shape = op), size = 3) + 
  scale_shape_manual(values = c("M", "N", "Q", "C")) +
  geom_errorbar(aes(ymin=YMin,ymax=YMax),width=0.1) +
  geom_line() +
  scale_y_continuous(limits = c(0,1)) +
  ylab("") +
  xlab("") +
  guides(shape = "none") +
  scale_colour_discrete(labels = c("Modal", "Negation", "Question", "Conditional")) +
  theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1)) +
  labs(title = "", color = "Operator")

@

\end{document}