### 2022-12-06 materials-lisa/analysis/3_projection-across-operators/data-analysis.Rnw

%!TEX TS-program = xelatex
%!TEX encoding = UTF-8 Unicode

%% document settings
\documentclass[10pt]{article}
\usepackage{euler}
\usepackage{xunicode,xltxtra}
\usepackage{fontspec}

% fonts
	\defaultfontfeatures{Mapping=tex-text, Ligatures=TeX}
	\setromanfont[Mapping=tex-text, Numbers={Proportional}]{Linux Biolinum}
	\setsansfont[Scale=MatchLowercase,Mapping=tex-text]{Optima}
	\setmonofont[Scale=MatchLowercase]{Andale Mono}

\usepackage[a4paper, margin = 1 in, left = .5 in, right = .5 in]{geometry}

\usepackage{setspace}
\usepackage[dvipsnames]{xcolor}

\usepackage{amsmath}
\usepackage{amssymb}
% \usepackage{mathtools}
\usepackage{linguex}
\usepackage{booktabs}

\usepackage[colorlinks=true,citecolor=MidnightBlue,urlcolor=MidnightBlue,linkcolor=MidnightBlue]{hyperref}

\title{Effects on projectivity ratings by Embedding Operator and Trigger --- \newline Data Analysis}
\author{Hofmann, deMarneffe, Tonhauser}
\date{\today}

\begin{document}
\setkeys{Gin}{width=1.0\textwidth}

<<setup, include=FALSE, echo=FALSE>>=
library(knitr)
hook_output = knit_hooks$get('output')
knit_hooks$set(output = function(x, options) {
  # this hook is used only when the linewidth option is not NULL
  if (!is.null(n <- options$linewidth)) {
    x = xfun::split_lines(x)
    # any lines wider than n should be wrapped
    if (any(nchar(x) > n)) x = strwrap(x, width = n)
    x = paste(x, collapse = '\n')
  }
  hook_output(x, options)
})

knitr::opts_chunk$set(echo=TRUE, prompt=TRUE, comment=NA, tidy.opts=list(width.cutoff=50), 
                      fig.path='figures/', fig.align = "center", out.width="\\linewidth")
options(tidy=TRUE, linewidth = 100)
@

\maketitle
\setcounter{tocdepth}{2}
\tableofcontents

\section{Introducing the dataset}
<<load-data, echo=FALSE, results='hide'>>=
# set working directory to directory of script
# this.dir <- dirname(rstudioapi::getSourceEditorContext()$path)
# setwd(this.dir)
setwd("/Users/lisa/Documents/research/GPP-experiments/materials-lisa/analysis/3_projection-across-operators")

# load data
df <- read.csv("../../../results/main/meta-analyses/projectivity/data/data_combined.csv", 
                    header=TRUE, sep=",")

@

<<data-overview>>=
str(df)
df$workerid <- as.factor(df$workerid)
length(levels(df$workerid))
@

\noindent The dataset consists of $57160$ observations from $2682$ participants (recruited on the online platforms Prolific and Amazon Mechanical Turk), across 12 experiments.

We are interested in how highly participants rate speaker commitment to the truth of an embedded complement clause, coded as \texttt{projective} on a real-numbered sliding scale between $0-1$.\\

\noindent The complement clause was embedded under an attitude verb, which in turn was embedded under an entailment-cancelling operator. Our fixed effects factors manipulate the following:

\begin{enumerate}
  \item The choice of attitude verb (coded as \texttt{verb})
  \item The entailment-cancelling operator (coded as \texttt{op})
\end{enumerate}

\noindent The levels for our fixed effects factors are the following:
<<fixed-efcts-lvls>>=
df$verb <- as.factor(df$verb)
levels(df$verb)
length(levels(df$verb))
df$op <- as.factor(df$op)
levels(df$op)
length(levels(df$op))
@

\noindent We are interested in the effect on \texttt{projective} of \texttt{verb} and \texttt{op}, as well as their interaction, corresponding to a $20 \times 4$ factorial design, yielding
<<ncond>>=
length(levels(df$verb))*length(levels(df$op))
@
conditions.\\

\noindent We have $20$ items, corresponding to the content of the complement clause.
<<nitems>>=
df$content <- as.factor(df$content)
levels(df$content)
length(levels(df$content))
@

\noindent We have roughly $36$ observations by item and condition. This is an approximate number, because the \texttt{op} manipulation is a between-studies manipulation, and the number of participants differs by experiment:
<<nobs>>=
# n observations
length(df[,1])

# observations by item
length(df[,1])/length(levels(df$content))
table(df$content)

# observations by verb
length(df[,1])/length(levels(df$verb))
table(df$verb)

# observations by operator
length(df[,1])/length(levels(df$op))
table(df$op)

# observations by item and condition
length(df[,1])/length(levels(df$content))/
  (length(levels(df$verb))*length(levels(df$op)))

@

\newpage
\section{Statistical Summaries and Graphs}
\subsection{Projectivity rating by operator}

Distribution of projectivity ratings by operator with means and $95\%$ bootstrapped confidence intervals.

<<summary-op, echo = FALSE, message=FALSE, cache=TRUE, out.width=".5\\textwidth">>=
# packages for data manipulation and plots
require(ggplot2)
require(dplyr)
require(forcats)

# load helper functions
source('../../../results/main/helpers.R')

pmeans_op = df %>% group_by(op) %>%
  summarize(Mean = mean(projective), CILow = ci.low(projective),
            CIHigh = ci.high(projective)) %>%
  mutate(YMin = Mean - CILow, YMax = Mean + CIHigh,
         op = fct_reorder(as.factor(op),Mean)) %>% ungroup()

pmeans_op %>% mutate(op = fct_reorder(op, Mean,
                                          .fun = mean)) %>%
  ggplot(aes(x = op, y=Mean)) +
  geom_violin(data=data.frame(op=df$op, Mean=df$projective),scale="width",color="gray80") +
  geom_point(size = 1, color = "lightblue") +
  geom_errorbar(aes(ymin=YMin,ymax=YMax), width=0.1, color = "lightblue") +
  geom_line(color = "lightblue")
  # labs(title = "Projectivity by operator")

@

\noindent The following generalizations emerge:
\begin{itemize}
  \item Conditionals have the highest projectivity ratings
  \item Projectivity ratings for questions are higher than those for modals and negation, but lower than those for conditionals
  \item Modals and negation have the lowest projectivity ratings
  \item The ratings for negation look a little higher than for modals, but error bars overlap
\end{itemize}

\noindent Although these differences appear to be significant, they are quite small.\\

\newpage
\subsection{Distributions of projectivity rating by verb with means and 95\% bootstrapped confidence interval:}

<<summary-verb, echo = FALSE, message=FALSE, cache=TRUE, fig.width=8, fig.height=4.5>>=

pmeans_verb = df %>% group_by(verb) %>%
  summarize(Mean = mean(projective), CILow = ci.low(projective), 
            CIHigh = ci.high(projective)) %>%
  mutate(YMin = Mean - CILow, YMax = Mean + CIHigh, 
         verb = fct_reorder(as.factor(verb),Mean))

# plot of projectivity means, 95% CIs and participants' ratings 
pmeans_verb %>% 
  ggplot(aes(x=fct_reorder(verb, Mean), y=Mean)) +
  geom_violin(data=data.frame(verb=df$verb, Mean=df$projective),scale="width",color="gray80") +
  geom_errorbar(aes(ymin=YMin,ymax=YMax),width=0.3,color="black") +
  geom_point(size=0.5,color="black") +
  scale_y_continuous(limits = c(0,1),breaks = c(0,0.2,0.4,0.6,0.8,1.0)) +
  theme(axis.text.x = element_text(size = 10, angle = 45, hjust = 1)) +
  xlab("") + ylab("")

@

\begin{itemize}
  \item We see gradual differences in projectivity between verbs
\end{itemize}

\newpage
\subsection{Distributions, means, and CIs for projectivity ratings by verb, for each operator:}
<<summary-combined-1, echo = FALSE, message=FALSE, cache=TRUE, fig.width=8, fig.height=6>>=
pmeans = df %>% group_by(verb, op) %>%
  summarize(Mean = mean(projective), CILow = ci.low(projective),
            CIHigh = ci.high(projective)) %>%
  mutate(YMin = Mean - CILow, YMax = Mean + CIHigh,
         verb = fct_reorder(as.factor(verb),Mean)) %>% ungroup()

pmeans %>%
  ggplot(aes(x=fct_reorder(verb, Mean), y=Mean)) +
  geom_violin(data=data.frame(verb=df$verb, Mean=df$projective, op=df$op),scale="width",color="gray80") +
  geom_errorbar(aes(ymin=YMin,ymax=YMax),width = 0.5, color="black") +
  geom_point(size=0.5,color="black") +
  facet_wrap(~op, labeller = labeller(op = c("m" = "Modal", "n" = "Negation", "q" = "Question", "c" = "Conditional"))) +
  scale_y_continuous(limits = c(0,1),breaks = c(0,0.2,0.4,0.6,0.8,1.0)) +
  theme(axis.text.x = element_text(size = 8, angle = 45, hjust = 1)) +
  ylab("") +
  xlab("")
@

\newpage
\subsection{Means and confidence intervals for projectivity rating by verb and operator:}
<<summary-combined-2, echo = FALSE, message=FALSE, cache=TRUE, fig.width=9, fig.height=4.7>>=

pmeans %>% 
  ggplot(aes(x=fct_reorder(verb, Mean), y=Mean, group = op, color = op)) +
  coord_cartesian(ylim=c(0,1)) +
  geom_point(aes(shape = op), size = 3) + 
  scale_shape_manual(values = c("C", "M", "N", "Q")) +
  geom_errorbar(aes(ymin=YMin,ymax=YMax),width=0.1) +
  geom_line() +
  scale_y_continuous(limits = c(0,1)) +
  ylab("") +
  xlab("") +
  # guides(shape = "none") +
  # scale_colour_discrete(labels = c("Modal", "Negation", "Question", "Conditional")) +
  theme_bw() +
  theme(legend.position = "none") +
  theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1))

@
\begin{itemize}
  \item We see interactions between verb and operator

  \item Two verbs show highest projectivity under negation: the anti-veridical \emph{pretend}, and the non-veridical \emph{think}. These are verb with relatively low overall projectivity.

  \item More projective verbs (\emph{\lq announce\rq} and above) have $\texttt{C} > \texttt{M}$

  \item Highly projective verbs (\emph{\lq hear \rq} and above) have  $\texttt{N} > \texttt{M}$

  \item We do not see any group of verbs that could be characterized as \lq semi-factive\rq\ in the sense of Karttunen.
  \begin{itemize}
    \item Specifically, \emph{discover} does not follow the predicted pattern: It is not more projective under negation, but most projective in conditionals and questions.
    \item For Karttunnen's \lq factives\rq, no difference between operators is expected.
    \item Kajsa Djärv about this distinction: cognitive predicates are semi-factive, and emotives are factive. The pattern suggested by karrtunen is also not found here (if anything, it might be the other way round)
  \end{itemize}

  \item For more generalizations, let's look at the same information plotted differently: By-operator projectivity for each verb
\end{itemize}

\newpage
\subsection{Projectivity ratings by operator, for each verb (Verb profiles)}
Version 1: With distributions of projectivity ratings as violin-plot
<<verb-profiles-1, echo = FALSE, message=FALSE, cache=TRUE, fig.width=14, fig.height=11>>=
pmeans %>%
  ggplot(aes(x = fct_reorder(op, Mean), y = Mean)) +
  coord_cartesian(ylim = c(0,1)) +
  facet_wrap(vars(fct_reorder(verb, Mean))) +
  geom_point(size = 1, color = "lightblue") +
  geom_violin(data=data.frame(verb=df$verb, Mean=df$projective, op=df$op),scale="width",color="gray80") +
  geom_errorbar(aes(ymin = YMin, ymax = YMax), width=0.1, color = "lightblue") +
  geom_line(color = "lightblue") +
  xlab("") + ylab("") +
  labs(title = "Mean projectivity by operator, for each verb")
@
Some more generalizations: We find different \lq profiles\rq\ for different verbs, of how embedding operators affect projectivity based on the verb. Groups of verbs show similar profiles:

\newpage
\noindent Version 2: With lines to show profiles
<<verb-profiles-2, echo = FALSE, message=FALSE, cache=TRUE, fig.width=14, fig.height=11>>=
pmeans %>%
  ggplot(aes(x = fct_reorder(op, Mean), y = Mean, group = verb)) +
  coord_cartesian(ylim = c(0,1)) +
  facet_wrap(vars(fct_reorder(verb, Mean))) +
  geom_point(size = 1, color = "lightblue") +
  geom_errorbar(aes(ymin = YMin, ymax = YMax), width=0.1, color = "lightblue") +
  geom_line(color = "lightblue") +
  xlab("") + ylab("") +
  theme_bw() +
  labs(title = "Mean projectivity by operator, for each verb")
@

\begin{itemize}
  \item \emph{pretend, think}: \textbf{anti-veridical profile}\\
    \texttt{N} $>$ \texttt{M, Q, C}, overall low projectivity

  \item \emph{acknowledge, hear, inform, see, discover, know, be annoyed}: \textbf{\lq factive\rq\ profile}\\
    \texttt{Q} $>$ \texttt{N, C} \textcolor{gray}{$>_?$} \texttt{M}, overall high projectivity (it may be possible to find further subgroups here)

  \item \emph{prove, confirm, establish, demonstrate, (announce), confess, admit, reveal}: \textbf{veridical profile}\\
    \texttt{M, C} $>$ \texttt{Q} $>$ \texttt{N}, overall med-lo to med-hi projectivity

  \item \emph{be right, suggest}: \textbf{reportative profile}\\
    \texttt{M, N, C} $>$ \texttt{Q}

\end{itemize}

Maybe these can have better names, not trying to suggest that verbs can neatly divided in factive v non-factive, but potentially this class / profile is what prompted intuitions in previous literature, and naming in this tradition could make sense, but can be changed depending on our rhetoric, of course.

\newpage

<<verb-profiles-3, echo = FALSE, message=FALSE, cache=TRUE, fig.width=14, fig.height=11>>=
vgroups <- data.frame(verb = levels(pmeans$verb), profile = 
                       c("0+0",  # acknowledge
                         "00+",   # admit 
                         "00+",  # announce
                         "-+-",  # be annoyed
                         "+-+",    # be right
                         "00+",   # confess
                         "+-+",   # confirm
                         "+0+",   # demonstrate
                         "0+0",  # discover
                         "+-+",   # establish
                         "-+0",  # hear
                         "-+-",  # inform
                         "-+-",  # know
                         "-00",    # pretend
                         "+-+",   # prove
                         "0++",   # reveal
                         "0-+",    # say
                         "-+-",  # see
                         "0-+",   # suggest
                         "-00"     # think
                         ))

lookup_group <- function(verb_name){
  return(as.character(filter(vgroups, verb == verb_name)[2]))
}

pmeans = pmeans %>% rowwise() %>% mutate(groups = lookup_group(verb))

pmeans %>%
  ggplot(aes(x = fct_reorder(op, Mean), y = Mean, group = verb, color = verb)) +
  coord_cartesian(ylim = c(0,1)) +
  facet_wrap(~groups) +
  geom_point(size = 1, color = "lightblue") +
  geom_errorbar(aes(ymin = YMin, ymax = YMax), width=0.1) +
  geom_line() +
  xlab("") + ylab("") +
  theme_bw() +
  labs(title = "Mean projectivity by operator, for each verb")
@

\end{document}