### 2023-11-17 results/main/13explicitIgnorance/models/analysis1/table1.tex
updated naturalness paper

 continue & \cellcolor{black} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} \\ 
  too & \cellcolor{white} & \cellcolor{black} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} \\ 
  also & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} \\ 
  again & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} & \cellcolor{yellow3} & \cellcolor{white} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} \\ 
  stop & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{black} & \cellcolor{white} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} \\ 
  \color{orange}be annoyed\color{black} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} \\ 
  cleft & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{black} & \cellcolor{white} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} & \cellcolor{yellow1} \\ 
  \color{orange}know\color{black} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{white} & \cellcolor{black} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow1} & \cellcolor{yellow1} \\ 
  \color{green}demonstrate\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{black} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} \\ 
  \color{green}pretend\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{black} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} \\ 
  \color{green}inform\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} \\ 
  \color{green}confess\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} \\ 
  \color{orange}see\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} \\ 
  \color{green}acknowledge\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} \\ 
  \color{orange}discover\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} \\ 
  \color{green}admit\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} \\ 
  \color{green}hear\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{black} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} \\ 
  \color{green}prove\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{white} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} \\ 
  \color{orange}reveal\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} & \cellcolor{yellow3} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} \\ 
  \color{green}announce\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{black} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} \\ 
  \color{green}be right\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} & \cellcolor{yellow2} \\ 
  \color{green}establish\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow3} & \cellcolor{yellow2} \\ 
  \color{green}confirm\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{black} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} \\ 
  \color{green}suggest\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{black} & \cellcolor{white} & \cellcolor{white} \\ 
  \color{green}think\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} & \cellcolor{white} \\ 
  \color{green}say\color{black} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{white} & \cellcolor{black} \\