### 2023-11-17 results/main/13explicitIgnorance/models/analysis2/table2.tex

 higher prior - EIC & \cellcolor{yellow1} & \cellcolor{yellow2} & \cellcolor{white} & \cellcolor{purple3} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{white} & \cellcolor{yellow3} & \cellcolor{white} & \cellcolor{white} & \cellcolor{purple2} & \cellcolor{white} & \cellcolor{white} & \cellcolor{white} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple3} & \cellcolor{purple2} \\ 
  lower prior - EIC & \cellcolor{yellow2} & \cellcolor{white} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple1} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} & \cellcolor{purple2} \\ 
  higher - lower prior & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{white} & \cellcolor{white} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow1} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} & \cellcolor{yellow2} \\