### 2024-02-05 results/main/meta-analyses/new/1_projectivity/projection-operators-summary.Rmd
added rmd overview file

---
title: "Projection operators data analysis summary"
output:
  pdf_document:
    toc: true
header-includes:
   - \usepackage[default]{sourcesanspro}
   - \usepackage{euler}
   - \usepackage[T1]{fontenc}
   - \usepackage[margin = 1in]{geometry}
   # - \usepackage{fontspec}
   # - \defaultfontfeatures{Mapping=tex-text, Ligatures=TeX}
   # - \setmonofont[Scale=MatchLowercase]{Andale Mono}
mainfont: SourceSansPro
date: "2024-01-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = 'center', tidy.opts=list(width.cutoff=50), 
                      fig.path='figures/', cache = TRUE)
library(tidyverse)

# summaries
source('../../../helpers.R')

# graphing
# Color blind friendly palette (http://www.cookbook-r.com/Graphs/Colors_(ggplot2)):
cbbPalette <- c("#E69F00", "#56B4E9", "#009E73", "#0072B2", "#F0E442", "#D55E00", "#CC79A7", "#000000")
cbbRotate <- c("#0072B2", "#CC79A7", "#000000", "#D55E00", "#F0E442", "#E69F00", "#56B4E9", "#009E73")
graphcolor <- "#0072B2"
theme_set(theme_bw())

# analysis
library(rstan)
options(mc.cores=parallel::detectCores())
rstan_options(auto_write=TRUE)
library(brms)
```

```{r load-data, include=FALSE}
this.dir <- dirname(rstudioapi::getSourceEditorContext()$path)
setwd(this.dir)
data <- read.csv("../data_combined.csv", header = TRUE, sep = ",")
data <- data %>% mutate(projection = as.numeric(projective), predicate = as.factor(verb), operator = as.factor(op), participant = as.factor(workerid), item = as.factor(content), .keep = "none")
levels(data$operator) <- c("conditional", "modal", "negation", "question")
```

## The dataset

The dataset consists of $57160$ observations from $2682$ participants (recruited on Amazon MTurk and Prolific). We are interested in how projection ratings (`projection`) are affected by `predicate` and `operator` .

```{r dataset}
str(data)
```

The levels of the independent variables are:

```{r independent-variables}
levels(data$operator)
levels(data$predicate)
```

\pagebreak

## Graphs and summaries

### Summary for overall projection ratings

```{r projection-means, echo=FALSE}
data %>% summarize(Mean = mean(projection), CILow = ci.low(projection), 
              CIHigh = ci.high(projection)) %>%
    mutate(YMin = Mean - CILow, YMax = Mean + CIHigh) %>% knitr::kable()
```

```{r projection-plot, echo=FALSE, fig.dim = c(4,2)}
data %>% ggplot(aes(x = projection)) +
  geom_area(stat = "bin", binwidth = .01, fill = graphcolor) +
  xlab("Certainty rating")
```

### Projection by operator

```{r operator-means, echo=FALSE}
operator_means <- data %>% group_by(operator) %>% 
    summarize(Mean = mean(projection), CILow = ci.low(projection), 
              CIHigh = ci.high(projection)) %>%
    mutate(YMin = Mean - CILow, YMax = Mean + CIHigh, 
           operator = fct_reorder(as.factor(operator), Mean)) %>% ungroup()
operator_means %>% knitr::kable()
operator_means <- operator_means %>% mutate(projection = Mean, .keep = "unused")
```

```{r operator-graph, echo=FALSE, fig.dim = c(3, 4.5), warning=FALSE, message=FALSE}
operator_means %>% ggplot(aes(x = fct_reorder(operator, projection, .fun = "mean"), 
                                y = projection, label = round(projection, digits = 2))) +
    geom_violin(data = data, scale="width", fill = graphcolor, 
                alpha = .4, color = graphcolor) +
    geom_errorbar(aes(ymin=YMin,ymax=YMax), width=0.1, color = "black") +
    geom_point(size=0.5, color = "black") +
    geom_text(hjust = 0, nudge_x = -0.08, nudge_y = -0.06) + 
    scale_y_continuous(limits = c(0,1), breaks = c(0,0.2,0.4,0.6,0.8,1.0),
                       labels = c("0","0.2","0.4","0.6","0.8","1"),
                       name = "Certainty rating") +
    theme(axis.text.x = element_text(size = 10, angle = 45, hjust = 1)) +
    xlab("Operator")
    # + labs(title = "Projection by operator")
```

### Projection by predicate

```{r predicate-means, echo=FALSE}
predicate_means <- data %>% group_by(predicate) %>% 
    summarize(Mean = mean(projection), CILow = ci.low(projection), 
              CIHigh = ci.high(projection)) %>%
    mutate(YMin = Mean - CILow, YMax = Mean + CIHigh, 
           predicate = fct_reorder(as.factor(predicate), Mean)) %>% ungroup()
predicate_means %>% knitr::kable()
predicate_means <- predicate_means %>% mutate(projection = Mean, .keep = "unused")

data$VeridicalityGroup = as.factor(
  ifelse(data$predicate %in% c("know", "discover", "reveal", "see", "be_annoyed"), "F", "NF"))

predicate_means$VeridicalityGroup = as.factor(
  ifelse(predicate_means$predicate %in% c("know", "discover", "reveal", "see", "be_annoyed"), "F", "NF"))
```

```{r color-coding, echo=FALSE}
colors <- c("darkorchid", "dodgerblue")
pred_order <- levels(fct_reorder(predicate_means$predicate, predicate_means$projection, .fun = "mean"))
textcolors <-  ifelse(pred_order  %in% c("know", "discover", "reveal", "see", "be_annoyed"), "darkorchid", "dodgerblue")
```

```{r predicate-graph, echo=FALSE, fig.dim = c(7, 4.5)}
predicate_means %>% ggplot(aes(x = fct_reorder(predicate, projection, .fun = "mean"), y = projection)) +
      geom_violin(data = data, scale="width", aes(fill = VeridicalityGroup), alpha = .4) +
      geom_errorbar(aes(ymin=YMin,ymax=YMax), width=0.3, color="black") +
      geom_point(size=2, aes(fill = VeridicalityGroup, shape = VeridicalityGroup), stroke=.5) +
      scale_x_discrete(name = "Predicate") +
      scale_y_continuous(limits = c(0,1), breaks = c(0,0.2,0.4,0.6,0.8,1.0),
                         labels = c("0","0.2","0.4","0.6","0.8","1"),
                         name = "Certainty rating") +
      scale_fill_manual(values = colors, name="Predicate type:", 
                          labels = c("(semi-)factives", "nonfactives")) +
      scale_shape_manual(values = c(21, 24, 25, 22, 23), name = "Predicate type:",
                         labels = c("(semi-)factives", "nonfactives")) +
      theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1, color = textcolors),
            panel.grid.major.x = element_blank(),
            legend.position="bottom", text = element_text(size=12)
            )
```

### Projection by predicate and operator

```{r predicate-operator-means, echo=FALSE, warning=FALSE, message = FALSE}
predicate_operator_means <- data %>% group_by(predicate, operator) %>%
      summarize(Mean = mean(projection), CILow = ci.low(projection), 
                CIHigh = ci.high(projection)) %>%
      mutate(YMin = Mean - CILow, YMax = Mean + CIHigh, 
             predicate = fct_reorder(as.factor(predicate),Mean)) %>%
      mutate(predicate = fct_reorder(predicate, Mean, .fun = mean)) %>%
      mutate(operator = fct_reorder(operator, Mean, .fun = mean)) %>% ungroup()
predicate_operator_means %>% knitr::kable()
predicate_operator_means <- predicate_operator_means %>% mutate(projection = Mean, .keep = "unused")

predicate_operator_means$VeridicalityGroup = as.factor(
  ifelse(predicate_operator_means$predicate %in% c("know", "discover", "reveal", "see", "be_annoyed"), "F", "NF"))
```

```{r predicate-operator-graph, echo=FALSE, warning=FALSE, fig.dim = c(7, 4.5)}
pred_order <- levels(fct_reorder(
  predicate_operator_means$predicate, 
  predicate_operator_means$projection, .fun = "mean"))
textcolors <-  ifelse(pred_order  %in% c("know", "discover", "reveal", "see", "be_annoyed"), "darkorchid", "dodgerblue")

predicate_operator_means %>%
  ggplot(aes(x = fct_reorder(predicate, projection), y = projection, group = operator, color = operator)) +
      geom_point(aes(shape = operator), size = 4) + 
      geom_errorbar(aes(ymin=YMin,ymax=YMax),width=0.1) +
      geom_line() +
      xlab("Predicate") +
      scale_y_continuous(limits = c(0,1), breaks = c(0,0.2,0.4,0.6,0.8,1), 
                         labels = c("0","0.2","0.4","0.6","0.8","1"),
                         name = "Mean certainty rating") +
      scale_shape_manual(values = c("M", "N", "C", "Q")) +
      scale_colour_manual(values=cbbPalette) +
      theme_bw() +
      theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1, color = textcolors),
            legend.position="none", text = element_text(size=12)
            )
```

```{r predicate-operator-graph-violin, echo=FALSE, fig.dim = c(10, 5)}
predicate_operator_means %>%
  ggplot(aes(x = fct_reorder(predicate, projection), y = projection)) +
    geom_violin(data = data, scale="width", fill = graphcolor, alpha = .4) +
    geom_errorbar(aes(ymin=YMin,ymax=YMax), width = 0.5, color="black") +
    geom_point(size=0.5, color="black") +
    xlab("Predicate") +
    scale_y_continuous(limits = c(0,1), breaks = c(0,0.2,0.4,0.6,0.8,1.0),
                       labels = c("0","0.2","0.4","0.6","0.8","1"),
                       name = "Mean certainty rating") +
    facet_wrap(~operator) +
    # theme(text = element_text(size=12), axis.text.x = element_text(size = 10, angle = 90, hjust = 1, vjust = -0.1, color=cols$Colors)) +
    theme(axis.text.x = element_text(size = 8, angle = 45, hjust = 1))
```

### Projection by predicate, negation only
```{r negation-predicate-graph, echo=FALSE, warning=FALSE, fig.dim = c(7, 3.2)}
pred_order <- levels(fct_reorder(
  subset(predicate_operator_means, operator == "negation")$predicate, 
  subset(predicate_operator_means, operator == "negation")$projection, .fun = "mean"))
textcolors <-  ifelse(pred_order  %in% c("know", "discover", "reveal", "see", "be_annoyed"), "darkorchid", "dodgerblue")

subset(predicate_operator_means, operator == "negation") %>%
  ggplot(aes(x = fct_reorder(predicate, projection, .fun = "mean"), y = projection)) +
      geom_violin(data = subset(data, operator == "negation"), scale="width", aes(fill = VeridicalityGroup), alpha = .4) +
      geom_errorbar(aes(ymin=YMin,ymax=YMax), width=0.3, color="black") +
      geom_point(size=2, aes(fill = VeridicalityGroup, shape = VeridicalityGroup), stroke=.5) +
      scale_x_discrete(name = NULL) +
      scale_y_continuous(limits = c(0,1), breaks = c(0,0.2,0.4,0.6,0.8,1.0),
                         labels = c("0","0.2","0.4","0.6","0.8","1"),
                         name = "Certainty rating") +
      scale_fill_manual(values = colors, name="Predicate type:", 
                          labels = c("(semi-)factives", "nonfactives")) +
      scale_shape_manual(values = c(21, 24, 25, 22, 23), name = "Predicate type:",
                         labels = c("(semi-)factives", "nonfactives")) +
      theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1, color = textcolors),
            panel.grid.major.x = element_blank(),
            legend.position="none", text = element_text(size=12)
            )

```

### Projection by predicate, question only
```{r question-predicate-graph, echo=FALSE, warning=FALSE, fig.dim = c(7, 3.2)}
pred_order <- levels(fct_reorder(
  subset(predicate_operator_means, operator == "question")$predicate, 
  subset(predicate_operator_means, operator == "question")$projection, .fun = "mean"))
textcolors <-  ifelse(pred_order  %in% c("know", "discover", "reveal", "see", "be_annoyed"), "darkorchid", "dodgerblue")

subset(predicate_operator_means, operator == "question") %>%
  ggplot(aes(x = fct_reorder(predicate, projection, .fun = "mean"), y = projection)) +
      geom_violin(data = subset(data, operator == "question"), scale="width", aes(fill = VeridicalityGroup), alpha = .4) +
      geom_errorbar(aes(ymin=YMin,ymax=YMax), width=0.3, color="black") +
      geom_point(size=2, aes(fill = VeridicalityGroup, shape = VeridicalityGroup), stroke=.5) +
      scale_x_discrete(name = NULL) +
      scale_y_continuous(limits = c(0,1), breaks = c(0,0.2,0.4,0.6,0.8,1.0),
                         labels = c("0","0.2","0.4","0.6","0.8","1"),
                         name = "Certainty rating") +
      scale_fill_manual(values = colors, name="Predicate type:", 
                          labels = c("(semi-)factives", "nonfactives")) +
      scale_shape_manual(values = c(21, 24, 25, 22, 23), name = "Predicate type:",
                         labels = c("(semi-)factives", "nonfactives")) +
      theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1, color = textcolors),
            panel.grid.major.x = element_blank(),
            legend.position="none", text = element_text(size=12)
            )
```

### Projection by predicate, modal only
```{r modal-predicate-graph, echo=FALSE, warning=FALSE, fig.dim = c(7, 3.2)}
pred_order <- levels(fct_reorder(
  subset(predicate_operator_means, operator == "modal")$predicate, 
  subset(predicate_operator_means, operator == "modal")$projection, .fun = "mean"))
textcolors <-  ifelse(pred_order  %in% c("know", "discover", "reveal", "see", "be_annoyed"), "darkorchid", "dodgerblue")

subset(predicate_operator_means, operator == "modal") %>%
  ggplot(aes(x = fct_reorder(predicate, projection, .fun = "mean"), y = projection)) +
      geom_violin(data = subset(data, operator == "modal"), scale="width", aes(fill = VeridicalityGroup), alpha = .4) +
      geom_errorbar(aes(ymin=YMin,ymax=YMax), width=0.3, color="black") +
      geom_point(size=2, aes(fill = VeridicalityGroup, shape = VeridicalityGroup), stroke=.5) +
      scale_x_discrete(name = NULL) +
      scale_y_continuous(limits = c(0,1), breaks = c(0,0.2,0.4,0.6,0.8,1.0),
                         labels = c("0","0.2","0.4","0.6","0.8","1"),
                         name = "Certainty rating") +
      scale_fill_manual(values = colors, name="Predicate type:", 
                          labels = c("(semi-)factives", "nonfactives")) +
      scale_shape_manual(values = c(21, 24, 25, 22, 23), name = "Predicate type:",
                         labels = c("(semi-)factives", "nonfactives")) +
      theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1, color = textcolors),
            panel.grid.major.x = element_blank(),
            legend.position="none", text = element_text(size=12)
            )
```

### Projection by predicate, conditional only
```{r conditional-predicate-graph, echo=FALSE, warning=FALSE, fig.dim = c(7, 3.75)}
pred_order <- levels(fct_reorder(
  subset(predicate_operator_means, operator == "conditional")$predicate, 
  subset(predicate_operator_means, operator == "conditional")$projection, .fun = "mean"))
textcolors <-  ifelse(pred_order  %in% c("know", "discover", "reveal", "see", "be_annoyed"), "darkorchid", "dodgerblue")

subset(predicate_operator_means, operator == "conditional") %>%
  ggplot(aes(x = fct_reorder(predicate, projection, .fun = "mean"), y = projection)) +
      geom_violin(data = subset(data, operator == "conditional"), scale="width", aes(fill = VeridicalityGroup), alpha = .4) +
      geom_errorbar(aes(ymin=YMin,ymax=YMax), width=0.3, color="black") +
      geom_point(size=2, aes(fill = VeridicalityGroup, shape = VeridicalityGroup), stroke=.5) +
      scale_x_discrete(name = NULL) +
      scale_y_continuous(limits = c(0,1), breaks = c(0,0.2,0.4,0.6,0.8,1.0),
                         labels = c("0","0.2","0.4","0.6","0.8","1"),
                         name = "Certainty rating") +
      scale_fill_manual(values = colors, name="Predicate type:", 
                          labels = c("(semi-)factives", "nonfactives")) +
      scale_shape_manual(values = c(21, 24, 25, 22, 23), name = "Predicate type:",
                         labels = c("(semi-)factives", "nonfactives")) +
      theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1, color = textcolors),
            panel.grid.major.x = element_blank(),
            legend.position="bottom", text = element_text(size=12)
            )
```

## Analysis

**Response:** projection ratings (rescale for beta regression). Scaling to transform from closed unit interval $[0,1]$ to open unit interval $(0,1)$, excluding boundaries using method used in Degen & Tonhauser (2022), from Smithson & Verkuilen (2006), for proportional data. $$y' = \frac{y \cdot (n - 1) + 0.5}{n}$$

```{r dependent-variable}
data$betaresponse <- (data$projection*(nrow(data)-1) + .5)/nrow(data)
data$betaresponse <- as.numeric(data$betaresponse)
summary(data$betaresponse)
str(data)
```

Fixed effect: operator. Using deviation coding to compare each level of operator to the overall mean.

```{r fixed-effect}
data <- data %>% mutate(operator = fct_reorder(operator, projection, .fun = "mean"))
contrasts(data$operator) <- contr.sum(4)
colnames(contrasts(data$operator)) <- c("modal", "negation", "question")
contrasts(data$operator)
```

## Predicate profiles

```{r predicate-profiles, echo=FALSE, fig.dim = c(10, 10)}
predicate_operator_means %>%
  ggplot(aes(x = fct_reorder(operator, projection, .fun = "mean"), y = projection, color = fct_reorder(operator, projection, .fun = "mean"), shape = fct_reorder(operator, projection, .fun = "mean"))) +
  geom_violin(data = data, scale="width", alpha = .4) +
  geom_errorbar(aes(ymin=YMin,ymax=YMax), width = 0.5, color="black", alpha = .6) +
  geom_point(size = 4) +
  facet_wrap(vars(fct_reorder(predicate, projection, .fun = "mean"))) +
  xlab("Operator") +
  scale_shape_manual(values = c("M", "N", "Q", "C")) +
  scale_colour_manual(values=cbbPalette) +
  scale_y_continuous(limits = c(0,1), breaks = c(0,0.2,0.4,0.6,0.8,1.0),
                       labels = c("0","0.2","0.4","0.6","0.8","1"),
                       name = "Certainty rating") +
  theme(axis.text.x = element_text(size = 8, angle = 45, hjust = 1), legend.position = "none")
```

Clustering predicates based on the differences between means for the operators.

```{r profiles}
# cross-tabulate means for predicate / operator
pomeans <- data.frame(matrix(nrow = length(levels(predicate_operator_means$predicate)),
                             ncol = length(levels(predicate_operator_means$operator))),
                      row.names = levels(predicate_operator_means$predicate))

colnames(pomeans) <- levels(predicate_operator_means$operator)

for (i in 1:length(levels(predicate_operator_means$predicate))) {
  current_pred = levels(predicate_operator_means$predicate)[i]
  for(j in 1:length(levels(predicate_operator_means$operator))) {
    current_op = levels(predicate_operator_means$operator)[j]
    pomeans[current_pred, current_op] <- subset(predicate_operator_means,
                                                predicate == current_pred &
                                                  operator == current_op)$projection[1]
  }
}

# get differences between operator means for each predicate
pomeansdiffs <- pomeans %>% mutate(mn = modal - negation, nc = negation - conditional, 
                                   cq = conditional - question, .keep = "none")

# plot hierarchical clusters based on differences
plot(hclust(dist(pomeansdiffs), method = "complete"))
```

### Negation high

```{r negation-high, echo=FALSE, fig.dim = c(2, 4)}
predicate_operator_means %>%  mutate(operator = fct_reorder(operator, projection,.fun = "mean")) %>%
  subset(predicate == "think" | predicate == "pretend") %>%
  ggplot(aes(x = operator, y = projection, group = predicate, color = predicate)) +
  geom_errorbar(aes(ymin=YMin,ymax=YMax), width = 0.2) +
  geom_point(size = 1) +
  geom_line() +
  annotate(geom = "text", x = 4, y = 0.3, label = "think", size = 2.8, color = "#CC79A7") +
  annotate(geom = "text", x = 3.85, y = 0.1, label = "pretend", size = 2.8, color = graphcolor) +
  scale_colour_manual(values=cbbRotate) +
  xlab("") +
  scale_y_continuous(limits = c(0,1), breaks = c(0,0.2,0.4,0.6,0.8,1.0),
                       labels = c("0","0.2","0.4","0.6","0.8","1"),
                       name = "Mean certainty rating") +
  theme(axis.text.x = element_text(size = 8, angle = 45, hjust = 1), legend.position = "none")
```

### Negation low

```{r negation-low, echo=FALSE, fig.dim = c(2, 4)}
predicate_operator_means %>% mutate(operator = fct_reorder(operator, projection,.fun = "mean")) %>% 
  subset(predicate == "confirm" | predicate == "prove" | predicate == "demonstrate"| predicate == "establish") %>%
  ggplot(aes(x = operator, y = projection, group = predicate, color = predicate)) +
  geom_errorbar(aes(ymin=YMin,ymax=YMax), width = 0.2) +
  geom_point(size = 1) +
  geom_line() +
  annotate(geom = "text", x = 3.4, y = 0.57, label = "demonstrate", size = 2.8, color = "#CC79A7") +
  annotate(geom = "text", x = 3.9, y = 0.28, label = "confirm", size = 2.8, color = graphcolor) +
  annotate(geom = "text", x = 2, y = 0.17, label = "prove", size = 2.8, color = "black") +
  scale_colour_manual(values=cbbRotate) +
  xlab("") +
  scale_y_continuous(limits = c(0,1), breaks = c(0,0.2,0.4,0.6,0.8,1.0),
                       labels = c("","","","","",""),
                       name = "") +
  theme(axis.text.x = element_text(size = 8, angle = 45, hjust = 1), legend.position = "none")
```

### Conditional high

```{r conditional-high, echo=FALSE, fig.dim = c(2, 4)}
predicate_operator_means %>% mutate(operator = fct_reorder(operator, projection,.fun = "mean")) %>% 
  subset(predicate == "announce" | predicate == "confess" | predicate == "admit" | predicate == "reveal") %>%
  ggplot(aes(x = operator, y = projection, group = predicate, color = predicate)) +
  geom_errorbar(aes(ymin=YMin,ymax=YMax), width = 0.2) +
  geom_point(size = 1) +
  geom_line() +
  annotate(geom = "text", x = 4, y = 0.79, label = "reveal", size = 2.8, color = "#D55E00") +
  annotate(geom = "text", x = 3.7, y = 0.48, label = "announce", size = 2.8, color = "#CC79A7") +
  annotate(geom = "text", x = 1, y = 0.64, label = "admit", size = 2.8, color = graphcolor) +
  annotate(geom = "text", x = 2, y = 0.44, label = "confess", size = 2.8, color = "black") +
  scale_colour_manual(values=cbbRotate) +
  xlab("Operator") +
  scale_y_continuous(limits = c(0,1), breaks = c(0,0.2,0.4,0.6,0.8,1.0),
                       labels = c("","","","","",""),
                       name = "") +
  theme(axis.text.x = element_text(size = 8, angle = 45, hjust = 1), legend.position = "none")
```

### Modal low

```{r modal-low, echo=FALSE, fig.dim = c(2, 4)}
predicate_operator_means %>% mutate(operator = fct_reorder(operator, projection,.fun = "mean")) %>% 
  subset(predicate == "know" | predicate == "inform" | predicate == "hear" | predicate == "be_annoyed") %>%
  ggplot(aes(x = operator, y = projection, group = predicate, color = predicate)) +
  geom_errorbar(aes(ymin=YMin,ymax=YMax), width = 0.2) +
  geom_point(size = 1) +
  geom_line() +
  annotate(geom = "text", x = 1, y = 0.45, label = "hear", size = 2.8, color = graphcolor) +
  annotate(geom = "text", x = 3, y = 0.92, label = "know", size = 2.8, color = "black") +
  annotate(geom = "text", x = 4, y = 0.68, label = "inform", size = 2.8, color = "#CC79A7") +
  scale_colour_manual(values=cbbRotate) +
  xlab("") +
  scale_y_continuous(limits = c(0,1), breaks = c(0,0.2,0.4,0.6,0.8,1.0),
                       labels = c("","","","","",""),
                       name = "") +
  theme(axis.text.x = element_text(size = 8, angle = 45, hjust = 1), legend.position = "none")
```

### Question high

```{r discover, echo=FALSE, fig.dim = c(2, 4)}
predicate_operator_means %>% mutate(operator = fct_reorder(operator, projection,.fun = "mean")) %>% 
  subset(predicate == "see" | predicate == "acknowledge" | predicate == "discover") %>%
  ggplot(aes(x = operator, y = projection, group = predicate, color = predicate)) +
  geom_errorbar(aes(ymin=YMin,ymax=YMax), width = 0.2) +
  geom_point(size = 1) +
  geom_line() +
  scale_colour_manual(values=cbbRotate) +
  annotate(geom = "text", x = 3.5, y = 0.63, label = "acknowledge", size = 2.8, color = graphcolor) +
  annotate(geom = "text", x = 1.2, y = 0.73, label = "discover", size = 2.8, color = "#CC79A7") +
  annotate(geom = "text", x = 3, y = 0.87, label = "see", size = 2.8, color = "black") +
  xlab("") +
  scale_y_continuous(limits = c(0,1), breaks = c(0,0.2,0.4,0.6,0.8,1.0),
                       labels = c("","","","","",""),
                       name = "") +
  theme(axis.text.x = element_text(size = 8, angle = 45, hjust = 1), legend.position = "none")
```

### Suggest

```{r suggest, echo=FALSE, fig.dim = c(2, 4)}
predicate_operator_means %>% mutate(operator = fct_reorder(operator, projection,.fun = "mean")) %>% 
  subset(predicate == "be_right" | predicate == "suggest") %>%
  ggplot(aes(x = operator, y = projection, group = predicate, color = predicate)) +
  geom_errorbar(aes(ymin=YMin,ymax=YMax), width = 0.2) +
  geom_point(size = 1) +
  geom_line() +
  scale_colour_manual(values=cbbRotate) +
  annotate(geom = "text", x = 3.5, y = 0.1, label = "be right", size = 2.8, color = graphcolor) +
  annotate(geom = "text", x = 3.9, y = 0.35, label = "suggest", size = 2.8, color = "#CC79A7") +
  xlab("Operator") +
  scale_y_continuous(limits = c(0,1), breaks = c(0,0.2,0.4,0.6,0.8,1.0),
                       labels = c("0","0.2","0.4","0.6","0.8","1"),
                       name = "") +
  theme(axis.text.x = element_text(size = 8, angle = 45, hjust = 1), legend.position = "none")
```

## R session info

```{r session-info, echo=FALSE, message=FALSE}
sessionInfo()
```